---
title: "Exercise_3"
author: "Zongyi Liu"
date: "2023-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Instruction in Original Excel

1.  Get 10 draws for each observation from N(0,1). Columns D to M. Note that this can be done before to save computation time.
2.  Compute values of equation 10 using the 10 draws. Columns O to X.
3.  Get 10 draws for alpha_t (equation 8). Columns Y to AH.
4.  Compute the likelihood given parms for each observation. Columns AI to AR.
5.  Now take the overall likelihood per physician by multiplying the likelihood for all t for each physician. The product is stored in the last row for each physician (the first 5 rows for each physician are not used).
6.  Find the average likelihood across all 10 observations (columns AS to BB) and then log (column BC).
7.  Multiply the average likelihood for each physician observation with last obs indicatore in column BD - same as column N - and put in column BE.
8.  Add up all the average likelihood across all physicians and put that in cell B6610 as the objective function.

## Practice

### 0_Read the Data for the Four Columns

```{r}
# install.packages("readxl")
# library(readxl)
# data_use<-read_excel('~./Users/ego/Desktop/897/ex 3/PhysDiff_Solve.xlsx')
```

```{r}
data_use<-read.csv("PhysDiff_Solve.csv")

date_use_2<-subset(data_use, select=c(Period.Pres,Upto.Pres,Cum.Det,Last.Obs))
```

### 1_Add Randomly Created Data

```{r cars}
c_1=rnorm(6000, mean=0, sd=1)
c_2=rnorm(6000, mean=0, sd=1)
c_3=rnorm(6000, mean=0, sd=1)
c_4=rnorm(6000, mean=0, sd=1)
c_5=rnorm(6000, mean=0, sd=1)
c_6=rnorm(6000, mean=0, sd=1)
c_7=rnorm(6000, mean=0, sd=1)
c_8=rnorm(6000, mean=0, sd=1)
c_9=rnorm(6000, mean=0, sd=1)
c_10=rnorm(6000, mean=0, sd=1)

dateset<-data.frame(
  Rnd_1=c_1,
  Rnd_2=c_2,
  Rnd_3=c_3,
  Rnd_4=c_4,
  Rnd_5=c_5,
  Rnd_6=c_6,
  Rnd_7=c_7,
  Rnd_8=c_8,
  Rnd_9=c_9,
  Rnd_10=c_10)

# Combine them together
data_use_4<-cbind(date_use_2,dateset)
data_use_4
```

### 2_Equation 10

```{r}
# =$B$6007*$C148+(SQRT($B$6009)*SQRT($C148))*M148

IQP = -1.29149813551524 #6006
TQP = 2.59183658947035  #6007
DSV	= 0.878787081       #6008
ADSV = exp(DSV)  # 6009

library(dplyr)

data_use_4 <- data_use_4%>%mutate(
  NDr_1 = TQP * Cum.Det + (sqrt(ADSV) * sqrt(Cum.Det)) * Rnd_1,
  NDr_2 = TQP * Cum.Det + (sqrt(ADSV) * sqrt(Cum.Det)) * Rnd_2,
  NDr_3 = TQP * Cum.Det + (sqrt(ADSV) * sqrt(Cum.Det)) * Rnd_3,
  NDr_4 = TQP * Cum.Det + (sqrt(ADSV) * sqrt(Cum.Det)) * Rnd_4,
  NDr_5 = TQP * Cum.Det + (sqrt(ADSV) * sqrt(Cum.Det)) * Rnd_5,
  NDr_6 = TQP * Cum.Det + (sqrt(ADSV) * sqrt(Cum.Det)) * Rnd_6,
  NDr_7 = TQP * Cum.Det + (sqrt(ADSV) * sqrt(Cum.Det)) * Rnd_7,
  NDr_8 = TQP * Cum.Det + (sqrt(ADSV) * sqrt(Cum.Det)) * Rnd_8,
  NDr_9 = TQP * Cum.Det + (sqrt(ADSV) * sqrt(Cum.Det)) * Rnd_9,
  NDr_10 = TQP * Cum.Det + (sqrt(ADSV) * sqrt(Cum.Det)) * Rnd_10)

data_use_4
```

### 3_Equation 8

```{r}
# (($B$6006)+(O2/$B$6009))/(1+($C2/$B$6009))
# O = NDr_1
# C = Rnd_1

data_use_4 <- data_use_4%>%mutate(
  qp_1 = (IQP+(NDr_1/ADSV))/(1+(Rnd_1/ADSV)),
  qp_2 = (IQP+(NDr_2/ADSV))/(1+(Rnd_2/ADSV)),
  qp_3 = (IQP+(NDr_3/ADSV))/(1+(Rnd_3/ADSV)),
  qp_4 = (IQP+(NDr_4/ADSV))/(1+(Rnd_4/ADSV)),
  qp_5 = (IQP+(NDr_5/ADSV))/(1+(Rnd_5/ADSV)),
  qp_6 = (IQP+(NDr_6/ADSV))/(1+(Rnd_6/ADSV)),
  qp_7 = (IQP+(NDr_7/ADSV))/(1+(Rnd_7/ADSV)),
  qp_8 = (IQP+(NDr_8/ADSV))/(1+(Rnd_8/ADSV)),
  qp_9 = (IQP+(NDr_9/ADSV))/(1+(Rnd_9/ADSV)),
  qp_10 = (IQP+(NDr_10/ADSV))/(1+(Rnd_10/ADSV)))

data_use_4
```

### 4_Likelihood Given Parms

```{r}
# ((EXP(Y6001)/(1+EXP(Y6001)))*A6001+(1/(1+EXP(Y6001)))*(1-A6001))^B6001
# Y = qp_1
# A = Period.Pres
# B = Upto.Pres

data_use_4 <- data_use_4%>%mutate(
  prp_1 = (exp(qp_1)/(1+exp(qp_1)))*Period.Pres + (1/(1+exp(qp_1))*(1-Period.Pres))^Upto.Pres,
  prp_2 = (exp(qp_2)/(1+exp(qp_2)))*Period.Pres + (1/(1+exp(qp_2))*(1-Period.Pres))^Upto.Pres,
  prp_3 = (exp(qp_3)/(1+exp(qp_3)))*Period.Pres + (1/(1+exp(qp_3))*(1-Period.Pres))^Upto.Pres,
  prp_4 = (exp(qp_4)/(1+exp(qp_4)))*Period.Pres + (1/(1+exp(qp_4))*(1-Period.Pres))^Upto.Pres,
  prp_5 = (exp(qp_5)/(1+exp(qp_5)))*Period.Pres + (1/(1+exp(qp_5))*(1-Period.Pres))^Upto.Pres,
  prp_6 = (exp(qp_6)/(1+exp(qp_6)))*Period.Pres + (1/(1+exp(qp_6))*(1-Period.Pres))^Upto.Pres,
  prp_7 = (exp(qp_7)/(1+exp(qp_7)))*Period.Pres + (1/(1+exp(qp_7))*(1-Period.Pres))^Upto.Pres,
  prp_8 = (exp(qp_1)/(1+exp(qp_8)))*Period.Pres + (1/(1+exp(qp_8))*(1-Period.Pres))^Upto.Pres,
  prp_9 = (exp(qp_1)/(1+exp(qp_9)))*Period.Pres + (1/(1+exp(qp_9))*(1-Period.Pres))^Upto.Pres,
  prp_10 = (exp(qp_1)/(1+exp(qp_10)))*Period.Pres + (1/(1+exp(qp_10))*(1-Period.Pres))^Upto.Pres
  )
  
data_use_4
```

### 6_Average Likelihood

```{r}
# Find the average likelihood across all 10 observations (columns AS to BB) and then log (column BC).
# =PRODUCT(AI2:AI7)
# =LN(AVERAGE(AS7:BB7))

data_use_5 <- data_use_4%>%mutate(
  ll_1  = lead(prp_1, 1)*lead(prp_1, 2)*lead(prp_1, 3)*lead(prp_1, 4)*lead(prp_1, 5)*lead(prp_1,6)
  ll_1  = lead(prp_1, 1)*lead(prp_1, 2)*lead(prp_1, 3)*lead(prp_1, 4)*lead(prp_1, 5)*lead(prp_1,6)
  
  ln(Avg(ll))=log(average(ll_1+ll_2+ll_3+ll_4+ll_5+ll_6+ll_7+ll_8+ll_9+ll_10))
  )

data_use_5
```

### 7_Average Plus Last Ob

```{r}
# Multiply the average likelihood for each physician observation with last obs indicatore in column BD - same as column N - and put in column BE.

data_use_6 <- data_use_5%>%mutate(
  BE_result=ln(Avg(ll))*Last.Obs
)
```

### 8_Add Up All

```{r}
LL=sum(data_use_6$BE_result)
```
